Precise predictions for same-sign W-boson scattering at the LHC

### 2018-03-05 Draft/conclusion.tex
Parton shower/parton-shower should eb consistent everywhere now+comments removed in conclusion

In the present article, a detailed study of the process $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$ at the LHC has been presented, 
mainly focused on the electroweak production mechanism which involves the scattering of massive vector-bosons.
Until very recently, when the complete calculation became available for the NLO QCD corrections (order $\mathcal O (\alphas\alpha^6)$),
the so-called VBS approximation was the standard for this kind of simulations. For this reason, various theoretical predictions 
have been compared to the full computation, both in a typical VBS fiducial region and also in more inclusive phase-space.
We have precisely quantified the differences that arise for several physical observables, 
in particular for the di-jet invariant mass and the rapidity-separation of the leading two jets.
This is the first time that such an in-depth study is performed.
%%%%and should be representative of the quality of the approximations for other VBS signatures.
Besides the study of fixed-order predictions, we have also investigated the impact of parton-shower.
To that end, several LO and NLO event generators 
which are able to perform matching to parton showers have been employed, and various observables have been thoroughly compared.
While in general observables which are described at NLO accuracy show reasonable agreement among the tools, larger differences can 
appear for those observables described at a lower accuracy, such as those that involve the third jet.
These differences appear in the central region where for VBS, and are the largest for those simulations which employ {\sc Pythia8}. The effect
has been understood, and it can be partially mitigated by changing the recoil scheme of the parton shower. These findings make it worth 
to further investigate these issues not only in the theoretical community, but also by experimental collaborations, for example by 
measuring related observables for similar processes.

@@

    \item In addition to the standard interpretation of EW signal versus QCD background, 
        combined measurements should also be presented as they are better defined theoretically. In fact, while at LO 
        the interference term can be included in the background component, at NLO the separation of EW and QCD component become more blurred, as, \emph{e.g.}\,
        at the order $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$ both types of amplitudes contribute.
        Therefore, a combined measurement including the EW, QCD, and interference contributions is desirable.
        \ZM{the following has to be clarified. one way would be to suggest to use  phase-space  cuts to enhance/suppress configurations 
            where EW/QCD are the dominant contributions}
Note that with such a measurement a comparison to the SM would
  be straigtforward and still be sensitive to the EW component.
  In addition the QCD component could be subtracted based on a
  well-defined Monte-Carlo prediction.
\MP{I am aware that this might be the most touchy point.
Therefore we should probably discuss it all together to ensure that we all agree on a definite statement.}

    \item Since the inclusion of NLO QCD corrections gives a better control of extra QCD radiations and reduces the ambiguities related to the 
        matching details and/or the parton-shower employed, we encourage the use of NLO-accurate event generators in experimental analyses.

>>>>

In the present article, a detailed study of the process $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$ at the LHC has been presented, 
mainly focused on the electroweak production mechanism which involves the scattering of massive vector-bosons.
Until very recently, when the complete calculation became available for the NLO QCD corrections (order $\mathcal O (\alphas\alpha^6)$),
the so-called VBS approximation was the standard for this kind of simulations. For this reason, various theoretical predictions 
have been compared to the full computation, both in a typical VBS fiducial region and also in more inclusive phase-space.
We have precisely quantified the differences that arise for several physical observables, 
in particular for the di-jet invariant mass and the rapidity-separation of the leading two jets.
This is the first time that such an in-depth study is performed.
%%%%and should be representative of the quality of the approximations for other VBS signatures.
Besides the study of fixed-order predictions, we have also investigated the impact of parton shower.
To that end, several LO and NLO event generators 
which are able to perform matching to parton showers have been employed, and various observables have been thoroughly compared.
While in general observables which are described at NLO accuracy show reasonable agreement among the tools, larger differences can 
appear for those observables described at a lower accuracy, such as those that involve the third jet.
These differences appear in the central region where for VBS, and are the largest for those simulations which employ {\sc Pythia8}. The effect
has been understood, and it can be partially mitigated by changing the recoil scheme of the parton shower. These findings make it worth 
to further investigate these issues not only in the theoretical community, but also by experimental collaborations, for example by 
measuring related observables for similar processes.

@@

    \item In addition to the standard interpretation of EW signal versus QCD background, 
        combined measurements should also be presented as they are better defined theoretically. In fact, while at LO 
        the interference term can be included in the background component, at NLO the separation of EW and QCD component become more blurred, as, \emph{e.g.}\,
        at the order $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$ both types of amplitudes contribute.
        Therefore, a combined measurement including the EW, QCD, and interference contributions is desirable.
        Note that with such a measurement a comparison to the SM would
        be straigtforward and still be sensitive to the EW component.
        In addition the QCD component could be subtracted based on a
        well-defined Monte-Carlo prediction.

    \item Since the inclusion of NLO QCD corrections gives a better control of extra QCD radiations and reduces the ambiguities related to the 
        matching details and/or the parton shower employed, we encourage the use of NLO-accurate event generators in experimental analyses.

### 2018-03-05 Draft/shower.tex
Parton shower/parton-shower should eb consistent everywhere now+comments removed in conclusion

We now discuss how different predictions compare when the matching to parton-shower is included. For such
a comparison we expect larger discrepancy than what we found at fixed-order, as a consequence of the different
matching schemes, parton shower employed, and of other details of the matching (such as the choice of the parton shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, the {\sc Powheg-Box}, and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. For {\sc VBFNLO} matched to {\sc Herwig} and the {\sc Powheg-Box}, we
restrict ourselves to showing results only in the VBS approximation,
{\emph i.e.}\ the $s$-channel contributions are neglected here. Besides,
also {\sc PHANTOM} is used for LO+PS results.\\
{\sc MG5\_aMC}, which
employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, will be used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 8.223)
and {\sc Herwig++}~\cite{Bahr:2008pv, Bellm:2013hwb} (version 2.7.1). The same parton showers will be employed for the LO results of {\sc PHANTOM}. For the {\sc Powheg-Box}, the namesake
matching procedure is employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} (version 8.230). {\sc VBFNLO} serves as a matrix-element and phase-space provider
for the {\sc Matchbox} module~\cite{Platzer:2011bc} of {\sc
Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx}, using an extended version of the BLHA
interface~\cite{Binoth:2010xt,Alioli:2013nda,Andersen:2014efa}. The {\sc Matchbox} module makes it
possible to choose between {\sc MC\-@NLO}-like and {\sc Powheg}-like
matching. As parton shower, both the default angular-ordered shower as
well as the dipole shower can be employed.
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected.

@@

In conclusion, the comparison of tools including matching with parton-shower clearly shows the benefits of the inclusion of NLO corrections: for most observables described
effectively at NLO accuracy differences between tools are at (or below) the $10\%$ level. 
Some exceptions exist, \emph{e.g.}\ the rapidity separation of the two tagging jets, which
on the one hand clearly suggest not to rely on a single tool/parton shower, and on the other make it worth to investigate more in details the way QCD radiation is
generated, \emph{e.g.}\ when fully-differential computations at NNLO will become available (for VBF Higgs production, see Refs.~\cite{Cacciari:2015jma, Cruz-Martinez:2018rod}). It is a remarkable fact that, even for those observables that display small discrepancies,
the theoretical uncertainty obtained via scale variations systematically underestimates the spread of predictions. Again, this stresses the need
 to employ at least two different tools in order to obtain a conservative estimate of theoretical uncertainties. Finally, the size of discrepancies for observables that are described at a lower perturbative accuracy, notably those related to the third jet, suggest that
experimental analyses should rely as little as possible on those observables and, in any case, use conservative estimate of the theory 
uncertainties. On the one hand, in order to improve the
description of these observables, a simulation of VBS+j at NLO accuracy, currently unavailable but within the reach of modern 
automated tools, is certainly desirable.
On the other hand, measurements of processes with similar 
color flow (EW production of a single vector boson plus jets,
VBF, \ldots) can certainly help in order to discriminate which tools perform better in the comparison with data~\cite{Aaboud:2017emo,Sirunyan:2017jej}.

>>>>

We now discuss how different predictions compare when the matching to parton shower is included. For such
a comparison we expect larger discrepancy than what we found at fixed-order, as a consequence of the different
matching schemes, parton shower employed, and of other details of the matching (such as the choice of the parton shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, the {\sc Powheg-Box}, and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. For {\sc VBFNLO} matched to {\sc Herwig} and the {\sc Powheg-Box}, we
restrict ourselves to showing results only in the VBS approximation,
{\emph i.e.}\ the $s$-channel contributions are neglected here. Besides,
also {\sc PHANTOM} is used for LO+PS results.\\
{\sc MG5\_aMC}, which
employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, will be used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 8.223)
and {\sc Herwig++}~\cite{Bahr:2008pv, Bellm:2013hwb} (version 2.7.1). The same parton showers will be employed for the LO results of {\sc PHANTOM}. For the {\sc Powheg-Box}, the namesake
matching procedure is employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} (version 8.230). {\sc VBFNLO} serves as a matrix-element and phase-space provider
for the {\sc Matchbox} module~\cite{Platzer:2011bc} of {\sc
Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx}, using an extended version of the BLHA
interface~\cite{Binoth:2010xt,Alioli:2013nda,Andersen:2014efa}. The {\sc Matchbox} module makes it
possible to choose between {\sc MC\-@NLO}-like and {\sc Powheg}-like
matching. As parton shower, both the default angular-ordered shower as
well as the dipole shower can be employed.
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected.

@@

In conclusion, the comparison of tools including matching to parton shower clearly shows the benefits of the inclusion of NLO corrections: for most observables described
effectively at NLO accuracy differences between tools are at (or below) the $10\%$ level. 
Some exceptions exist, \emph{e.g.}\ the rapidity separation of the two tagging jets, which
on the one hand clearly suggest not to rely on a single tool/parton shower, and on the other make it worth to investigate more in details the way QCD radiation is
generated, \emph{e.g.}\ when fully-differential computations at NNLO will become available (for VBF Higgs production, see Refs.~\cite{Cacciari:2015jma, Cruz-Martinez:2018rod}). It is a remarkable fact that, even for those observables that display small discrepancies,
the theoretical uncertainty obtained via scale variations systematically underestimates the spread of predictions. Again, this stresses the need
 to employ at least two different tools in order to obtain a conservative estimate of theoretical uncertainties. Finally, the size of discrepancies for observables that are described at a lower perturbative accuracy, notably those related to the third jet, suggest that
experimental analyses should rely as little as possible on those observables and, in any case, use conservative estimate of the theory 
uncertainties. On the one hand, in order to improve the
description of these observables, a simulation of VBS+j at NLO accuracy, currently unavailable but within the reach of modern 
automated tools, is certainly desirable.
On the other hand, measurements of processes with similar 
color flow (EW production of a single vector boson plus jets,
VBF, \ldots) can certainly help in order to discriminate which tools perform better in the comparison with data~\cite{Aaboud:2017emo,Sirunyan:2017jej}.
